---
title: "Biostats Project EDA"
author: "Alexander Ruse"
date: "November 26, 2024"
output: html_document
---

```{r}
library(tidyverse)
library(RSocrata)
df <- read.socrata(
  url = "https://data.cdc.gov/resource/n8mc-b4w4.json?$limit=50000"
)
```

```{r}
head(df)
```


Creating a crosstabs for 
```{r}
# Load required libraries
library(tidyverse)

# Assuming your dataset is called `df`, create crosstabs for all pairs of variables
variables <- c("hosp_yn", "symptom_status", "exposure_yn", "ethnicity", "age_group", "race", "sex")

# Generate crosstabs using base R's table()
for (i in 1:(length(variables) - 1)) {
  for (j in (i + 1):length(variables)) {
    var1 <- variables[i]
    var2 <- variables[j]
    cat("\nCrosstab of", var1, "and", var2, ":\n")
    print(table(df[[var1]], df[[var2]], useNA = "ifany"))
  }
}


```
Summary of results: The "0 - 17 years" group has a higher count in Missing and Unknown, suggesting incomplete data for this group. The Yes row has relatively low counts, indicating fewer individuals were hospitalized compared to No or Unknown. Hospitalization likelihood increases with age, a trend consistent with many health outcomes. For "0 - 17 years," 5594 entries are Missing and 3047 are Unknown. Incomplete data might bias the results or reduce the reliability of conclusions for these groups.
 

Interpretation of results: 
hosp_yn (Hospitalization Outcome) strongly varies with age_group, symptom_status, and race.
Other factors like ethnicity and exposure_yn may have additional effects and missing data (e.g., NA, Unknown) is prevalent and needs careful handling.

Significant Patterns:
symptom_status x exposure_yn: Symptomatic individuals show higher hospitalization likelihood in combination with exposure.
age_group x race: Older age groups exhibit variable hospitalization patterns across racial categories.
ethnicity x race: Ethnic background interacts with race categories in influencing outcomes.
sex x race: Sex disparities are evident across races

Potential Interaction Terms:
Based on the crosstabs, we are considering pursuing the following interaction terms

1. age_group x symptom_status
Why: Age affects the likelihood of hospitalization, but this effect may differ depending on whether an individual is symptomatic or asymptomatic.
Example: Older symptomatic individuals may have a much higher risk of hospitalization compared to younger symptomatic individuals.

2. age_group x race
Why: Racial disparities in health outcomes often vary across age groups. Older individuals in certain racial categories might have higher hospitalization risks.
Example: Interaction between age group and race could highlight compounded vulnerabilities.

3. age_group x ethnicity
Why: Ethnicity impacts health outcomes, but the degree of impact may vary by age.
Example: Hospitalization differences in Hispanic versus Non-Hispanic populations across age groups.

4. symptom_status x exposure_yn
Why: Exposure and symptom status are closely related to hospitalization risk.
Example: Symptomatic individuals with known exposures likely have a higher probability of hospitalization.

5. race x ethnicity
Why: Ethnic background and race jointly contribute to disparities in health outcomes.
Example: Certain racial groups within Hispanic populations may have distinct hospitalization patterns.

6. sex x race
Why: Gender differences in hospitalization outcomes can vary by race.
Example: Female versus male hospitalization likelihood might differ across racial groups.

7. age_group x sex
Why: Hospitalization differences between sexes could vary with age.
Example: Older women might have different risk profiles compared to older men.

Chi-squared testing done to illustrate interaction terms
```{r}
chisq.test(table(df$hosp_yn, df$age_group))

```
The result of the Chi-squared test indicates that there is a statistically significant association between hosp_yn (hospitalization outcome) and age_group as the p-value is extremely small and considered well below the standard level of significance (0.05)

```{r}
ggplot(df, aes(x = age_group, fill = hosp_yn)) +
  geom_bar(position = "fill") +
  labs(y = "Proportion", title = "Hospitalization Outcomes by Age Group")

```

now let's try different chi-squared tests from our data to see which interaction terms we can find:
```{r}
chisq.test(table(df$age_group, df$sex))

```

```{r}
chisq.test(table(df$race, df$ethnicity))
```

```{r}
df_recode <- df |>
  mutate(across(everything(), ~if_else(. %in% c("Unknown", "Missing", "NA"), NA_character_, .))) |>
  mutate(hosp_yn = factor(hosp_yn, levels = c("No", "Yes")))

df_recode$age_group <- ordered(df_recode$age_group)
```

```{r}
df_recode
```

```{r}
library(brms)
```


```{r}
mod_1 <- brm(
  hosp_yn ~ age_group + race,
  data = df_recode,
  iter = 1000,
  chains = 4,
  cores = getOption("mc.cores", 4),
  family = bernoulli
)
```

```{r}
summary(mod_1)
```


```{r}
library(tidyverse)
library(tidybayes)
library(posterior)
library(tidybayes)
library(bayesplot) 
```

```{r}
pp_check(mod_1, type = "stat", bins = 30)
```

