---
title: "LW_Missing"
author: "Luke Wilsen"
date: "2024-11-26"
output: html_document
---

# Notebook Idea: What to do with missing data, and why?

```{r}
load("/Users/lukewilsen/Desktop/School/BS_730/BS_730_proj/730_data_150000_obs.RData")
df
```

# Investigating missingness

```{r}
dfna = df %>% 
  mutate(across(everything(), ~ na_if(.x, "Missing") %>% na_if("Unknown") %>% na_if("NA")))
```

```{r}
colSums(is.na(dfna))
```

There is no Month, State, county, and current status data missing from the dataset.

```{r}
#is.na(dfna %>% select(underlying_conditions_yn, hosp_yn, age_group))
```


## hosp_yn Samples per state, age group, underlying_condition_yn, symptoms_yn

```{r}
dfna %>% 
  select(age_group, 
        sex, 
        race, 
        ethnicity, 
        exposure_yn, 
        symptom_status, 
        underlying_conditions_yn,
        hosp_yn) %>% 
  tbl_summary(by = hosp_yn)
```

```{r}
print(c("% of rows without known hospitalization status:", as.character(sum(is.na(dfna$hosp_yn)) / sum(length(dfna$hosp_yn)))))
```

```{r}
colnames(dfna)

interest_cols = c("sex","race","ethnicity","hosp_yn","underlying_conditions_yn","age_group")

for (col in interest_cols){
  print(ggplot(dfna)+
          geom_bar(aes(x = hosp_yn, fill = !!sym(col)), position = "dodge")
        )
}
```

## Methods for dealing with missing data:

- https://pmc.ncbi.nlm.nih.gov/articles/PMC11101000/ 


## Method 1: Deletion

```{r}
dfNoNa = drop_na(dfna)
```

```{r}
for (col in interest_cols){
  print(ggplot(dfNoNa)+
          geom_bar(aes(x = hosp_yn, fill = !!sym(col)), position = "fill")
        )
}
```

## Initial results

- Increased proportion of females hospitalized, compared to females not hospitalized
- Increased proportion of Blacks hospitalized, compared to Blacks not hospitalized
  - Decreased proportion of Whites hospitalized, compared to Whites not hospitalized
- 0-17 yrs seems to be protective, 65+ years looks like it significantly increases risk of hospitalization
  - 18-49 seems slightly protective
