---
title: "Models"
author: "Luke Wilsen"
date: "2024-11-26"
output: html_document
---

```{r}
library(skimr)
library(tidyverse)
library(RSocrata)
library(gtsummary)
library(brms)
library(modelr)
library(bayesplot)
library(rstanarm)

load("/Users/lukewilsen/Desktop/School/BS_730/BS_730_proj/730_data_150000_obs.RData")
```

```{r}
df

dfna = df %>% 
  mutate(across(everything(), ~ na_if(.x, "Missing") %>% na_if("Unknown") %>% na_if("NA")))
```

```{r}
logit <- function(p) log(p/(1-p))

invlogit <- function(x) 1/(1+exp(-x))
```


# Initial considerations/concerns
- Our explanatory variables are mostly (if not entirely) categorical
  - What kind of recoding scheme should we use?
    - One-hot encoding
    - Recode age_group to be an ordinal variable (with scores to represent distances between the groups)
  
- Are any of the variables we're interested in correlated?
  - Yes (all of them are correlated/collinear)
- What priors?
  - Should we incorporate prior information?
  - Jeffrey's prior bc logistic regression?
  
# Recoding
```{r}
dfna$black = ifelse(dfna$race == "Black", 1, 0)
dfna$white = ifelse(dfna$race == "White", 1, 0)
dfna$asian = ifelse(dfna$race == "Asian", 1, 0)
dfna$mult = ifelse(dfna$race == "Multiple/Other", 1, 0)
dfna$anai = ifelse(dfna$race == "American Indian/Alaska Native", 1, 0)

age_scores1 = c(8.5, 33.5, 57, 75)
# 75 might be too low of a representation of the 65+ group, so we'll try another model with a higher value

age_scores2 = c(8.5, 33.5, 57, 80)

dfna <- dfna %>%
  mutate(age_scores1 = case_when(
    age_group == "0 - 17 years" ~ age_scores1[1],
    age_group == "18 to 49 years" ~ age_scores1[2],
    age_group == "50 to 64 years" ~ age_scores1[3],
    age_group == "65+ years" ~ age_scores1[4],
    is.na(age_group) ~ NA_real_  # Handle NA explicitly
  ))

dfna <- dfna %>%
  mutate(age_scores2 = case_when(
    age_group == "0 - 17 years" ~ age_scores2[1],
    age_group == "18 to 49 years" ~ age_scores2[2],
    age_group == "50 to 64 years" ~ age_scores2[3],
    age_group == "65+ years" ~ age_scores2[4],
    is.na(age_group) ~ NA_real_  # Handle NA explicitly
  ))

dfna = dfna %>% 
  mutate(hosp_yn = case_when(
    hosp_yn == "No" ~ 0,
    hosp_yn == "Yes" ~ 1
  ))
```

```{r}
select_df = dfna %>% select(age_group, race, res_state, hosp_yn) %>% drop_na()
```

# Models
## basemod

Model specs:
- Base priors
- logist link function
```{r results = "hide"}
basemod <- brm(
  hosp_yn ~ age_group + race,
  data = dfna,
  iter = 1500,
  chains = 4,
  cores = getOption("mc.cores", 4),
  family = bernoulli
)
```

### interpret
```{r}
summary(basemod)
```

Base group of the model: American Indians/Alaska Natives in the 0-17 years age group.
- Get odds of hospitalization for this group

Compared to 0-17 yr old American Indians/Alaska Natives, the odds of White people being hospitalized are (1 - exp(-0.67)) = 0.4882914 = 48.83% lower.

Compared to 0-17 yr old American Indians/Alaska Natives, the odds of Black people being hospitalized are (1 - exp(-0.07)) = 6.76% lower.

Compared to 0-17 yr old American Indians/Alaska Natives, the odds of Asian people being hospitalized are (1 - exp(-0.26)) = 22.89% lower.

Compared to 0-17 yr old American Indians/Alaska Natives, the odds of Asian people being hospitalized are (1 - exp(-0.26)) = 42.88% lower.

Increasing age ==> Higher odds of hospitalization

```{r}
1 - exp(-0.67)
1 - exp(-0.07)
1 - exp(-0.26)
1 - exp(-0.56)
unique(dfna$age_group)
unique(dfna$race)
```

### mod check: basemod

```{r}

vars = c("Intercept","b_age_group18to49years","b_age_group50to64years","b_age_group65Pyears","b_raceAsian","b_raceBlack","b_raceWhite")

vars2 = c("Intercept","b_age_group18to49years","b_age_group50to64years","b_age_group65Pyears", "b_black", "b_white", "b_asian", "b_mult", "b_anai")

for (var in vars){
  plot(basemod, variable = var)
}
```

Pretty good mixing of the chains, Rhat = 1 for all params, min Bulk ESS = 1572, min Tail ESS = 1588
 - Overall I feel fairly confident that this model fits the data well.
  - However, I want to see how changing the coding of age group affects this.
  
  
### Post pred check

```{r}
pp_check(basemod, type = "stat")
```

#### State by state post pred check
```{r}
select_i_1 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(1,5)]))
select_i_2 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(6,10)]))
select_i_3 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(11,15)]))
select_i_4 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(16,20)]))
select_i_5 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(21,25)]))
select_i_6 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(26,30)]))
select_i_7 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(31,35)]))
select_i_8 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(36,40)]))
select_i_9 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(41,45)]))
select_i_10 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(46,48)]))

ynew_bmod = posterior_predict(basemod)

color_scheme_set("blue")
ppc_group_bmod = function(select_i){
  return(ppc_stat_grouped(select_df$hosp_yn[select_i], yrep = ynew_bmod[, select_i],
                 group = select_df$res_state[select_i], 
                 stat = "mean") + theme_classic())
}

ppc_group_bmod(select_i_1)
ppc_group_bmod(select_i_2)
ppc_group_bmod(select_i_3)
ppc_group_bmod(select_i_4)
ppc_group_bmod(select_i_5)
ppc_group_bmod(select_i_6)
ppc_group_bmod(select_i_7)
ppc_group_bmod(select_i_8)
ppc_group_bmod(select_i_9)
ppc_group_bmod(select_i_10)
```

## encoded1 mod
```{r, results='hide', echo=FALSE}
encmod1 <- brm(
  hosp_yn ~ age_scores1 + race,
  data = dfna,
  iter = 1500,
  chains = 4,
  cores = getOption("mc.cores", 4),
  family = bernoulli
)
```

```{r}
summary(encmod1)
```

### mod check
```{r}
vars = c("Intercept","b_age_group18to49years","b_age_group50to64years","b_age_group65Pyears","b_raceAsian","b_raceBlack","b_raceWhite")

vars2 = c("Intercept","b_age_group18to49years","b_age_group50to64years","b_age_group65Pyears", "b_black", "b_white", "b_asian", "b_mult", "b_anai")

vars3 = c("Intercept","b_age_scores1", "b_raceAsian","b_raceBlack","b_raceWhite", "b_raceMultipleDOther")

for (var in vars3){
  plot(encmod1, variable = var)
}
```

```{r}
pp_check(encmod1, type = "stat")
conditional_effects(encmod1)
```





## encoded2 mod
Changed which age scoring used
```{r, results='hide', echo=FALSE}
encmod2 <- brm(
  hosp_yn ~ age_scores2 + race,
  data = dfna,
  iter = 1500,
  chains = 4,
  cores = getOption("mc.cores", 4),
  family = bernoulli
)
```

```{r}
summary(encmod2)
```

```{r}
vars4 = c("Intercept","b_age_scores2", "b_raceAsian","b_raceBlack","b_raceWhite", "b_raceMultipleDOther")


for (var in vars4){
  plot(encmod2, variable = var)
}
```

```{r}
pp_check(encmod2, type = "stat")
conditional_effects(encmod2)
```

## basemod hierarchical




```{r}
bmheirarchical <- brm(
  hosp_yn ~ age_group + race + (1 | res_state),
  data = select_df,
  iter = 2500,
  chains = 4,
  cores = getOption("mc.cores", 4),
  family = bernoulli
)
```

```{r}
summary(bmheirarchical)
```

### Model checking
```{r}
pp_check(bmheirarchical, type = "stat")
conditional_effects(encmod1)
```

```{r}
obs_hosp <- dfna |>
  select(age_group, race, res_state, hosp_yn) |>
  na.omit() |>
  reframe(prop = sum(hosp_yn == 1) / n(), .by = "res_state")

# Show a plot by state, with predictions and observed values
dfna |>
  data_grid(res_state, age_group, race) |>
  add_epred_draws(bmheirarchical, allow_new_levels = TRUE) |>
  ggplot(aes(x = .epred, y = res_state)) +
    stat_pointinterval() +
    scale_y_discrete(limits = rev) +
    geom_point(data = obs_hosp, aes(x = prop, y = res_state), color = "green")+
  labs(title = "Prediction interval for proportion hospitalized vs observed proportion")
```


## encmod2 hierarchical
```{r}
encmod2heirarchical <- brm(
  hosp_yn ~ age_scores2 + race + (1 | res_state),
  data = dfna,
  iter = 2500,
  chains = 4,
  cores = getOption("mc.cores", 4),
  family = bernoulli
)
```

```{r}
summary(encmod2heirarchical)
```

### Model checking
```{r}
dfna |>
  data_grid(res_state, age_scores2, race) |>
  add_epred_draws(encmod2heirarchical, allow_new_levels = TRUE) |>
  ggplot(aes(x = .epred, y = res_state)) +
    stat_pointinterval() +
    scale_y_discrete(limits = rev) +
    geom_point(data = obs_hosp, aes(x = prop, y = res_state), color = "green")+
  labs(title = "Prediction interval for proportion hospitalized vs observed proportion")
```

```{r}
pp_check(encmod2heirarchical, type = "stat")

select_i_1 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(1,5)]))
select_i_2 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(6,10)]))
select_i_3 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(11,15)]))
select_i_4 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(16,20)]))
select_i_5 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(21,25)]))
select_i_6 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(26,30)]))
select_i_7 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(31,35)]))
select_i_8 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(36,40)]))
select_i_9 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(41,45)]))
select_i_10 <- which(is.element(select_df$res_state,
                               unique(select_df$res_state)[seq(46,48)]))

ynew_si = posterior_predict(encmod2heirarchical)

color_scheme_set("blue")
ppc_group = function(select_i){
  return(ppc_stat_grouped(select_df$hosp_yn[select_i], yrep = ynew_si[, select_i],
                 group = select_df$res_state[select_i], 
                 stat = "mean") + theme_classic())
}

ppc_group(select_i_1)
ppc_group(select_i_2)
ppc_group(select_i_3)
ppc_group(select_i_4)
ppc_group(select_i_5)
ppc_group(select_i_6)
ppc_group(select_i_7)
ppc_group(select_i_8)
ppc_group(select_i_9)
ppc_group(select_i_10)
```




